% The first command in your LaTeX source must be the \documentclass command.
\documentclass[]{ceurart}

\sloppy
\usepackage{listings}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\lstset{breaklines=true}

\begin{document}

\copyrightyear{2026}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

\conference{CLEF 2026}

\title{Job-Skill Matching with Skill Type Classification}

\tnotemark[1]
\tnotetext[1]{Working notes paper submitted to CLEF 2026.}

\author[1]{Hussain Mustansir}[
email=hm08436@st.habib.edu.pk,
]
\cormark[1]
\fnmark[1]
\author[1]{Mahrukh Yousuf}[
email=my08055@st.habib.edu.pk,
]
\fnmark[1]
\author[1]{Afiya Sohail}[
email=as09186@st.habib.edu.pk,
]
\fnmark[1]
\address[1]{Habib University, Karachi, Pakistan}

\cortext[1]{Corresponding author.}
\fntext[1]{These authors contributed equally.}

\begin{abstract}
This literature review addresses the task of Job-Skill Matching with Skill Type
Classification. The task requires, given a job title, retrieving a ranked list of relevant
skills and classifying each as core, complementary, or transversal. We are provided a
training set of at least 5,000 job title--skill pairs, a development set of 200 job titles
with ESCO-normalised skill annotations carrying a core/contextual relevance label, and a
test set of 500 job titles for blind evaluation. We survey multiple prior works covering 
skill classification and skill extraction, each contributing methods or findings we draw on
in building our approach.
\end{abstract}

\begin{keywords}
  job-skill matching \sep
  skill type classification \sep
  ESCO \sep
  transformer models \sep
  named entity recognition \sep
  ensemble learning \sep
  information retrieval
\end{keywords}

\maketitle

%% ---------------------------------------------------------------
\section{Introduction to the Task}
%% ---------------------------------------------------------------

The task of Job-Skill Matching with Skill Type Classification sits at the intersection of
information retrieval and multi-class labelling. A system must take a raw job title as
input, retrieve a ranked list of skills from the ESCO taxonomy, and simultaneously assign
each retrieved skill one of three labels: \textit{core}, \textit{complementary}, or
\textit{transversal}. The challenge lies in combining semantic retrieval with
skill-type classification under ranking-based evaluation metrics such as nDCG and MAP.
Because TalentCLEF includes multilingual data, models must operate across languages
without relying solely on translation. 

%% ---------------------------------------------------------------
\section{Literature Review}
%% ---------------------------------------------------------------

The work most relevant to our task spans two directions: retrieving skills from a given job
title and classifying skills by type using taxonomy-aligned labels. The papers
reviewed below each tackle a different piece of this problem, yet together they outline
the methodological foundations we build upon.

The study titled ``Techniques for Transversal Skill Classification and Relevant Keyword
Extraction from Job Advertisements''~\cite{Gavrilescu2025}
presents a two-stage classification pipeline that identifies transversal skills in job
ads according to the ESCO taxonomy. The first stage places a skill into one of six
broad groups---Core, Thinking, Self-management, Social/communication, Physical, and
Life skills---while the second stage refines that assignment into finer subclasses.
Both stages use TF-IDF word features as inputs to fully connected neural network
classifiers. In addition to classification, the authors extract keywords associated
with each skill category using feature importance metrics and TF-IDF scoring. The
six ESCO transversal categories used in this work provide a useful reference for how
skill types can be structured, which is relevant to the type-level labelling required
by TalentCLEF.

% Building on the theme of skill extraction from unstructured text, the work titled
% ``Job Description Parsing with Explainable Transformer Based Ensemble Models to Extract
% the Technical and Non-Technical Skills''~\cite{Akkasi2024} takes an ensemble
% route that combines five extractors. The author scraped 30,695 job advertisements from LinkedIn,
% Glassdoor, Indeed, and StepStone covering 22 IT roles, and annotated them using a
% curated inventory of 7,170 hard skills and 269 soft skills. The resulting multi-task
% ensemble reached F-scores of 67\% for non-technical skills and 72\% for technical
% skills. Since TalentCLEF evaluates systems using ranking metrics, combining multiple
% model outputs through voting can help produce more stable results. The fact that
% soft skills are harder to extract than hard skills also hints at a similar challenge
% in our task, where transversal skills are often described in less precise wording
% than core skills.

Another contribution is the work titled ``SkillSpan: Hard and Soft Skill
Extraction from English Job Postings'' ~\cite{Zhang2022} which introduces 
\textit{JobBERT} and \textit{JobSpanBERT},
domain-adapted BERT variants pre-trained on 3.2 million unlabelled job-posting
sentences. JobBERT achieved the highest combined span-F1 of 59.73 on the test set. 
For our project: domain-adaptive 
pre-training on job-posting text will be a good option, which
motivates a similar strategy for TalentCLEF; and skill spans, particularly for soft
skills, are substantially longer than typical named entities, so any extraction
component must be evaluated on multi-token spans rather than single-token accuracy.

Moving from skill extraction in job advertisements to the candidate side of the
matching problem, the work titled ``A Two-Phase Framework for Automated Information
Extraction from Curriculum Vitae'' ~\cite{Ngo2020CVExtraction} proposes
an end-to-end pipeline for pulling structured fields out of CVs. The framework
operates in two stages: a layout-analysis-based block segmentation step that partitions
each CV into regions such as Personal Information, Education, Work Experience, and
Skills, followed by a named entity recognition model that labels tokens within each
region using IOB notation. The authors compare three
transformer-based models: XLM-RoBERTa, BERT Multilingual Cased, and BERT Base
Uncased on a dataset of 914 English and 400 Vietnamese CVs covering
seven entity types including skill sets. XLM-RoBERTa achieved the best overall F1 of
approximately 80\%, though multi-word entities such as degree
names, organisation names, and job titles remain harder to detect than shorter fields
like phone numbers or emails.

Shifting the perspective from span-level extraction to document-level prediction, the
study titled ``Retrieving Skills from Job Descriptions: A Language Model Based Extreme
Multi-label Classification Framework'' ~\cite{Bhola2020XMLC} frames
skill prediction as an extreme multi-label classification (XMLC) problem over roughly
2,500 skill labels from a Singaporean government job portal. Their BERT-XMLC model
encodes a job description through BERT\textsubscript{BASE}, compresses the
representation via a bottleneck layer, and predicts each skill independently through
sigmoid outputs. They also introduced \textit{Correlation Aware Bootstrapping (CAB)}, 
which expands the training data by using patterns in how skills appear together 
and by encoding skill names. As a result, the model’s performance improved, increasing 
Recall@5 by about 10\% and nDCG@5 by about 7\% compared to the original model. 

Focusing specifically on the transversal dimension of skills, the work titled
``Hierarchical Classification of Transversal Skills in Job Advertisements Based on
Sentence Embeddings'' ~\cite{Leon2024Hierarchical} develops a two-level
classification pipeline aligned with the ESCO transversal-skill hierarchy. At the
first level, six binary classifiers determine whether a sentence refers to one of the
six ESCO transversal categories; at the second level, a multi-label classifier assigns
specific sub-skills within each category. All models use 768-dimensional BERT-based
sentence embeddings. The authors built a dataset of 5,208 manually labelled sentences
from 219 European job ads. The multilingual embedding model performed almost the same 
as the English-only model, with only about a 0.55\% difference in performance. This 
supports the result reported by Gavrilescu et al.~\cite{Gavrilescu2025}, suggesting 
that translating job advertisements into English may not be necessary when working with 
multilingual European data.

Another work titled “Comparing Discriminative, Contrastive, and Prompt-Based Methods for 
Job Title and Skill Matching” ~\cite{zhang2025nlpnorth} has focused on automatically identifying the most relevant 
skills associated with a given job title Their work features three approaches: classification based, contrastive learning, 
and prompt based embedding methods.  For the job to skill matching task, the classification based 
approach reformulates the ranking problem as a binary classification task. A multilingual 
encoder generates embeddings for job titles and skills, and a classification layer on top of 
it predicts if the skill is relevant, allowing the model to learn job-skill relationships from 
labeled data. During training, each job title and correct skill pair was treated as a positive 
example, while incorrect job-skill pairs were generated as negative 
examples by randomly sampling unrelated skills. In contrast, the contrastive learning approach 
trains embedding models to place related job titles and skills closer in the vector space, while 
prompt-based models rely on pretrained, instruction-tuned embedding models in a zero-shot setting 
to rank skills based on cosine similarity. The results showed that classification based models 
performed best for the job title to skill matching task. The highest MAP score achieved by 
classification models was 0.290, outperforming the other two approaches. 

The next study “Beyond Titles: Semantic Matching of Jobs and Skills Using LLMs and S-BERT” ~\cite{vazquez2025beyond}
focuses on the same problem from TalentCLEF 2025 Task B, i.e. to predict which skills match 
a given job title. Instead of simply comparing the job title with skill names, the authors 
tried to understand the meaning behind them. Their method first uses large language models 
to generate short descriptions of both the job titles and the skills. These descriptions 
give more context about what the job or skill actually involves. Then they convert these 
descriptions into dense vector embeddings using a Sentence-BERT model; all-MiniLM-L6-v2. 
Once both jobs and skills are represented as vectors, cosine similarity is used to measure 
the semantic relatedness between job and skill representations.To improve reliability, 
the authors used multiple prompt setups and models to generate the descriptions, 
combining all results to produce a final ranking of the most relevant skills for each 
job title. The proposed ensemble model achieved a MAP score of 0.2442 on the development 
set and 0.278 on the test set, outperforming the official baseline system which relied 
only on direct title embeddings. The results show how adding meaningful descriptions for 
jobs and skills can help improve job title matching with the right skills. 

Similarly, a study titled “Enhancing Job-Skill Matching with LLM-Driven Data Augmentation 
and Fine-Tuned Bi-Encoders” ~\cite{ali2025jobskill} also focuses on the problem in TalentCLEF 2025 Task B challenge. 
Initially, they experimented with two types of models: cross-encoders and bi-encoders. 
Cross-encoders allow the model to understand job title and skill relationship more deeply, 
but they are very slow. While bi-encoders are much faster since they create embeddings for 
jobs and skills independently but perform poorly as job titles and skill names are very short 
and lack context. To solve this, the authors enrich the input data by adding descriptions, a 
large language model was used to generate descriptions for each job title. The final system 
uses a fine-tuned bi-encoder model that compares job and skill embeddings using cosine 
similarity. The models were trained using contrastive learning so that relevant job-skill 
pairs receive higher similarity scores than unrelated ones. After experimenting with several 
pretrained models, the best results were achieved using the E5-Instruct encoder combined 
with LLM-generated job descriptions and ESCO skill descriptions. The proposed system achieved 
a MAP score of 0.345 on the official test set, showing that adding contextual information 
through LLM-generated descriptions can significantly improve job-skill matching performance. 

Finally, the work titled ``Unsupervised Skill Identification from Job Ads'' 
~\cite{Khaouja2021Unsupervised} takes a different route by avoiding
labelled training data entirely. The methodology has two stages: first, job ad
sentences are embedded alongside entries from the ESCO skill base using Sent2vec,
and the closest ESCO skill is retrieved for each sentence via cosine similarity;
second, capitalised words not captured in the first stage are looked up against
Wikipedia and DBpedia to identify programming languages, software names, and
certifications. ESCO occupation clusters are then used to filter out spurious
matches based on skill co-occurrence within the same ISCO-08 group. Evaluated on
100 manually labelled job ads from diverse sectors, the Sent2vec variant achieved
a precision of 83\% and a recall of 78\% for hard skills, improving recall by 56\%
over a simple exact-match baseline. Notably, SBERT embeddings performed
considerably worse in this setting, suggesting that off-the-shelf sentence
transformers may need domain-specific fine-tuning for skill retrieval. 

\begin{table*}[ht]
\centering
\caption{Summary of reviewed literature.}
\label{tab:summary}
\small
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}p{3.5cm}p{2.5cm}p{2.8cm}p{3.0cm}p{3.2cm}@{}}
\toprule
\textbf{Paper Title} & \textbf{Task} & \textbf{Dataset} & \textbf{Model(s)} & \textbf{Key Result} \\
\midrule
Techniques for Transversal Skill Classification~\cite{Gavrilescu2025}
  & Transversal skill classification
  & ESCO-aligned job ads
  & BERT Sentence Transformer + neural classifiers
  & Multilingual within 0.55\% of English-only accuracy \\
\addlinespace
% Job Description Parsing with Ensemble Models~\cite{Akkasi2024}
%   & Hard \& soft skill extraction (NER)
%   & 30,695 scraped IT job ads; 22 roles; 7,170+ skill forms
%   & Ensemble of BiLSTM, CNN, CRF with BERT embeddings; LIME
%   & F1 67\% (soft), 72\% (hard); significant over baselines \\
% \addlinespace
SkillSpan: Hard and Soft Skill Extraction~\cite{Zhang2022}
  & Span-level skill \& knowledge extraction
  & 391 JPs, 14.5K sentences, 12.5K spans
  & JobBERT, JobSpanBERT (domain-adapted BERT); CRF layer
  & Span-F1 59.73 (combined); STL $>$ MTL \\
\addlinespace
A Two-Phase Framework for CV Information Extraction~\cite{Ngo2020CVExtraction}
  & CV information extraction (NER + layout segmentation)
  & 914 English + 400 Vietnamese CVs; 7 entity types
  & XLM-RoBERTa, BERT Multilingual, BERT Base; CRF layer
  & F1 80\% (XLM-RoBERTa); segmentation 97\% accuracy \\
\addlinespace
Retrieving Skills from Job Descriptions: BERT-XMLC~\cite{Bhola2020XMLC}
  & Skill prediction as extreme multi-label classification
  & 20,298 JPs; 2,548 skill labels
  & BERT-XMLC with bottleneck layer; CAB bootstrapping
  & Recall@5 +10\%; nDCG@5 +7\%; 86\% skills implicit \\
\addlinespace
Hierarchical Classification of Transversal Skills~\cite{Leon2024Hierarchical}
  & Hierarchical transversal skill classification
  & 219 EURES job ads; 5,208 labelled sentences
  & Neural classifiers on BERT sentence embeddings; two-level hierarchy
  & Level-1 accuracy $>$94\%; level-2 $\sim$68\% \\

\addlinespace
Unsupervised Skill Identification from Job Ads~\cite{Khaouja2021Unsupervised}
  & Unsupervised skill identification
  & 100 labelled job ads; ESCO skill base
  & Sent2vec + ESCO similarity retrieval; Wikipedia/DBpedia
  & Precision 83\%, Recall 78\% (hard skills) \\

\addlinespace
Enhancing Job-Skill Matching with LLM-Driven Data Augmentation and Fine-Tuned Bi-Encoders~\cite{ali2025jobskill}
  & Job–skill matching
  & TalentCLEF 2025 Task B dataset based on the ESCO skills database
  & Pretrained encoders (ESCOXLM-R, E5-Large, E5-Instruct)
  & Best system achieved MAP = 0.345 on the TalentCLEF test set \\

\addlinespace
Comparing Discriminative, Contrastive, and Prompt-Based Methods for Job Title and Skill Matching
~\cite{zhang2025nlpnorth}
  & Job–skill matching
  & ESCO-based job titles and skills dataset used in TalentCLEF 2025 Task B
  & mDeBERTa-v3-base (classification), multilingual-e5-large (contrastive), gte-Qwen2-7B-instruct (prompt-based)
  & Best MAP: 0.290 achieved by classification model\\

\addlinespace
Beyond Titles: Semantic Matching of Jobs and Skills Using LLMs and S-BERT
~\cite{vazquez2025beyond}
  & Job–skill matching
  & ESCO-based job titles and skills dataset used in TalentCLEF 2025 Task B
  & S-BERT: all-MiniLM-L6-v2; LLMs for definition generation: gemma-3-4b-it-qat, gemma3:27b
  & Ensemble MAP: 0.2442 (Dev), 0.278 (Test); Baseline MAP: 0.1874 (Dev), 0.196 (Test)\\
  
\bottomrule
\end{tabular}
\end{table*}

%%
%% Bibliography
\bibliographystyle{IEEEtran}
\bibliography{bibfile}

%%
\appendix

\end{document}